Methodology for Assessing the State of the Practice for Domain X

### 2020-11-10 StateOfPractice/Methodology/Methodology.tex
updated section 9, methodology
Reformatted methodology document to have fixed width for lines of text

\section{How to Identify the Domain} \label{SecIdentifyDomain}
A domain of research software must be identified. Research software is defined in this exercise as ``software that is used to generate, process or analyze results that [are intended] to appear in a publication'' \citep{hettrick2014uk}. The chosen domain must have the following properties:

@@

\section{How to Identify Candidate Software} \label{SecIdentifyCandSoft}
The candidate software can be found through search engine queries targeting authoritative lists of software from the domain, such as lists on GitHub and swMATH, and by searching domain related publications. Domain experts are also asked for their suggestions and are asked to review the list. The candidate software should have the following properties:

@@

\section{How to Initially Filter the Software List} \label{SecInitialFilter}
If the list of software is too long (over around 30 packages), then steps need to be taken to create a more manageable list. To reduce the length of the list, the following filters are applied in the priority order listed, stopping once the desired number of projects is reached:

@@

\section{Empirical Measures} \label{SecEmpiricalMeasures}
Some quality measurements rely on the gathering of raw and processed empirical data. We focus on data that is reasonably easy to collect, which we combine and analyze for this assessment. The measures that are collected relate to the research questions that we want to answer. For instance, we collect some of the data to see how large a project is. Other measures intend to ascertain a project’s popularity, or how active the project is.
Section \ref{rawdata} will orient the reader as to what raw data is collected. 
Some of this data can be observed from Git repository metrics. The rest can be collected using freeware tools. \href{https://github.com/tomgi/git_stats}{GitStats} is used to measure the number of binary files as well as the number of added and deleted lines in a repository. The tool is also used to measure the number of commits over different intervals of time. \href{https://github.com/boyter/scc}{Sloc Cloc and Code (scc)} is used to measure the number of text based files as well as the number of total, code, comment, and blank lines in a repository. These tools were selected due to their installability, usability, and ability to gather the empirical measures listed below. A guide for installing and running them can be found \href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/A Guide to Empirical Measures.pdf}{here}. Section \ref{processeddata} introduces the required processed data, which is calculated using the raw data.

\subsection{Raw Data}\label{rawdata}
The following raw data measures are extracted from repositories:

@@

\subsection{Processed Data}\label{processeddata}
The following measures are calculated from the raw data:

@@

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=1.0\textwidth]{measurement_template}
		\caption{Top of Measurement Template}
		\label{measurement_template_image}

	\end{center}
\end{figure}

The \href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Combined_MeasurementTemplate_EmpiricalMeasures.xlsx}{Measurement Template} is used to track measurements and quality scores for all of the software packages in the domain. For each software package, fill out one column of the template. This process can take between 2 to 5 hours for each package.
Project developers can be contacted for help regarding installation, if necessary, but a cap of about 2 hours should be imposed on the entire installation process.

@@

\begin{enumerate}
	\item Publications about the software: A list of publications can be found directly on the website of some software packages. For others use Google Scholar or a similar index.
	\item Initial release date: Mark the release year if an exact date is not available.
	\item Continuous integration: Search the software artifacts for any mention of continuous integration. The search function on GitHub can help. 
	\item Is there evidence that performance was considered?: Search the software artifacts for any mention of speed, storage, throughput, performance optimization, parallelism, multi-core processing, or similar considerations. The search function on GitHub can help.
	\item Getting started tutorial: Sometimes this is found within another artifact, like the user manual.
\end{enumerate}

@@

\begin{enumerate} 
	\item Gather the summary information into the top section of the document (Figure \ref{measurement_template_image})
	\item Using the GitStats tool that is described in Section \ref{SecEmpiricalMeasures} gather the measurements for the Repo Metrics - GitStats section found near the bottom of the document
	\item Using the SCC tool that is also described in Section \ref{SecEmpiricalMeasures} gather the measurements for the Repo Metrics - SCC section found near the bottom of the document
	\item If the software package is found on git, gather the measurements for the Repo Metrics - GitHub section found near the bottom of the document
	\item Review installation documentation and attempt to install the software package on a virtual machine
	\item Gather the measurements for installability
	\item Gather the measurements for correctness and verifiability
	\item Gather the measurements for surface reliability
	\item Gather the measurements for surface robustness
	\item Gather the measurements for surface usability
	\item Gather the measurements for maintainability
	\item Gather the measurements for reusability
	\item Gather the measurements for surface understandability
	\item Gather the measurements for visibility and transparency
	\item Assign a score out of ten for each quality. The score can be measured using the \href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/MeasurementTemplate_ImpressionCalculator.xlsx}{Measurement Template Impression Calculator}. For each quality measurement, the file indicates the appropriate score to assign the measurement based on possible measurement values.
\end{enumerate}

\section{Analytic Hierarchy Process} \label{SecAHP}
The Analytical Hierarchy Process (AHP) is a decision-making technique that can be used when comparing multiple options by multiple criteria. In our work AHP is used for comparing and ranking the software packages of a domain using the quality scores that are gathered in the \href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Combined_MeasurementTemplate_EmpiricalMeasures.xlsx}{Measurement Template}. AHP performs a pairwise analysis between each of the quality options using a matrix which is then used to generate an overall score for each software package for the given criteria. \cite{SmithEtAl2016} shows how AHP is applied to ranking software based on quality measures. We have developed a tool for conducting this process. The tool includes an AHP JAR script and a sensitivity analysis JAR script that is used to ensure that the software package rankings are appropriate with respect to the uncertainty of the quality scores. The README file of the tool is found \href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/AHP2020/LBM/README.txt}{here}. This file outlines the requirements for, and configuration and usage of, the JAR scripts. The JAR scripts, source code, and required libraries are located in the same folder as the README file.

@@

An AHP pairwise comparison between the short list packages with respect to usability and modifiability survey results is conducted by the domain expert. The tool can be found here. The usability and modifiability survey can be found \href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Experiments.pdf}{here}. This step is intended to start a conversation about the features of specific software packages.

@@

\appendix
\section{Appendix}
\subsection{Survey for the Selected Projects}
\ad{Several questions are borrowed from \href{https://gitlab.cas.mcmaster.ca/smiths/pub/-/blob/master/Jegatheesan2016.pdf}{Jegatheesan2016}, and needed to be cited later.}
\subsubsection{Information about the developers and users}
\begin{enumerate}
\item Interviewees' current position/title? degrees?
\item Interviewees' contribution to/relationship with the software?
\item Length of time the interviewee has been involved with this software?
\item How large is the development group?
\item What is the typical background of a developer?
\item How large is the user group?
\item What is the typical background of a user?
\end{enumerate}

@@

\subsection{Empirical Measures Considerations - Processed Data}
Metrics that can be calculated from the raw data.

@@

\subsection{Empirical Measures Considerations - Tool Tests}
\ad{This section is currently a note of unorganized contents. Most parts will beremoved or relocated.}

@@

\subsubsection{git-stats}
\href{https://github.com/tomgi/git_stats}{GitHub repo}

@@

\subsubsection{scc}
\href{https://github.com/boyter/scc}{GitHub repo}

\subsubsection{git-of-theseus}
\href{https://github.com/erikbern/git-of-theseus}{GitHub repo}

@@

\subsubsection{hercules}
\href{https://github.com/src-d/hercules}{GitHub repo}

@@

\subsubsection{gitinspector}
\href{https://github.com/ejwa/gitinspector}{GitHub repo}

>>>>

\section{How to Identify the Domain} \label{SecIdentifyDomain}

A domain of research software must be identified. Research software is defined
in this exercise as ``software that is used to generate, process or analyze
results that [are intended] to appear in a publication''
\citep{hettrick2014uk}. The chosen domain must have the following properties:

@@

\section{How to Identify Candidate Software} \label{SecIdentifyCandSoft}

The candidate software can be found through search engine queries targeting
authoritative lists of software from the domain, such as lists on GitHub and
swMATH, and by searching domain related publications. Domain experts are also
asked for their suggestions and are asked to review the list. The candidate
software should have the following properties:

@@

\section{How to Initially Filter the Software List} \label{SecInitialFilter}

If the list of software is too long (over around 30 packages), then steps need
to be taken to create a more manageable list. To reduce the length of the list,
the following filters are applied in the priority order listed, stopping once
the desired number of projects is reached:

@@

\section{Empirical Measures} \label{SecEmpiricalMeasures}

Some quality measurements rely on the gathering of raw and processed empirical
data. We focus on data that is reasonably easy to collect, which we combine and
analyze for this assessment. The measures that are collected relate to the
research questions that we want to answer. For instance, we collect some of the
data to see how large a project is. Other measures intend to ascertain a
project’s popularity, or how active the project is.  Section \ref{rawdata} will
orient the reader as to what raw data is collected.  Some of this data can be
observed from Git repository metrics. The rest can be collected using freeware
tools. \href{https://github.com/tomgi/git_stats}{GitStats} is used to measure
the number of binary files as well as the number of added and deleted lines in a
repository. The tool is also used to measure the number of commits over
different intervals of time. \href{https://github.com/boyter/scc}{Sloc Cloc and
  Code (scc)} is used to measure the number of text based files as well as the
number of total, code, comment, and blank lines in a repository. These tools
were selected due to their installability, usability, and ability to gather the
empirical measures listed below. A guide for installing and running them can be
found
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/A
  Guide to Empirical Measures.pdf} {here}. Section \ref{processeddata}
introduces the required processed data, which is calculated using the raw data.

\subsection{Raw Data}\label{rawdata}

The following raw data measures are extracted from repositories:

@@

\subsection{Processed Data}\label{processeddata}

The following measures are calculated from the raw data:

@@

\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=1.0\textwidth]{measurement_template}
    \caption{Top of Measurement Template}
    \label{measurement_template_image}
  \end{center}
\end{figure}

The
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Combined_MeasurementTemplate_EmpiricalMeasures.xlsx}
{Measurement Template} is used to track measurements and quality scores for all
of the software packages in the domain. For each software package, fill out one
column of the template. This process can take between 2 to 5 hours for each
package.  Project developers can be contacted for help regarding installation,
if necessary, but a cap of about 2 hours should be imposed on the entire
installation process.

@@

\begin{enumerate}
\item Initial release date: Mark the release year if an exact date is not
  available.
\item Publications about the software: A list of publications can be found
  directly on the website of some software packages. For others use Google
  Scholar or a similar index.
\item Is there evidence that performance was considered?: Search the software
  artifacts for any mention of speed, storage, throughput, performance
  optimization, parallelism, multi-core processing, or similar
  considerations. The search function on GitHub can help.
\item Getting started tutorial: Sometimes this is found within another artifact,
  like the user manual.
\item Continuous integration: Search the software artifacts for any mention of
  continuous integration. The search function on GitHub can help.
\end{enumerate}

@@

\begin{enumerate}
\item Gather the summary information into the top section of the document
  (Figure \ref{measurement_template_image})
\item Using the GitStats tool that is described in Section
  \ref{SecEmpiricalMeasures} gather the measurements for the Repo Metrics -
  GitStats section found near the bottom of the document
\item Using the SCC tool that is also described in Section
  \ref{SecEmpiricalMeasures} gather the measurements for the Repo Metrics - SCC
  section found near the bottom of the document
\item If the software package is found on git, gather the measurements for the
  Repo Metrics - GitHub section found near the bottom of the document
\item Review installation documentation and attempt to install the software
  package on a virtual machine
\item Gather the measurements for installability
\item Gather the measurements for correctness and verifiability
\item Gather the measurements for surface reliability
\item Gather the measurements for surface robustness
\item Gather the measurements for surface usability
\item Gather the measurements for maintainability
\item Gather the measurements for reusability
\item Gather the measurements for surface understandability
\item Gather the measurements for visibility and transparency
\item Assign a score out of ten for each quality. The score can be measured
  using the
  \href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/MeasurementTemplate_ImpressionCalculator.xlsx}
  {Measurement Template Impression Calculator}. For each quality measurement,
  the file indicates the appropriate score to assign the measurement based on
  possible measurement values.
\end{enumerate}

\section{Analytic Hierarchy Process} \label{SecAHP}

The Analytical Hierarchy Process (AHP) is a decision-making technique that can
be used when comparing multiple options by multiple criteria. In our work AHP is
used for comparing and ranking the software packages of a domain using the
quality scores that are gathered in the
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Combined_MeasurementTemplate_EmpiricalMeasures.xlsx}
{Measurement Template}. AHP performs a pairwise analysis between each of the
quality options using a matrix which is then used to generate an overall score
for each software package for the given criteria. \cite{SmithEtAl2016} shows how
AHP is applied to ranking software based on quality measures. We have developed
a tool for conducting this process. The tool includes an AHP JAR script and a
sensitivity analysis JAR script that is used to ensure that the software package
rankings are appropriate with respect to the uncertainty of the quality
scores. The README file of the tool is found
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/AHP2020/LBM/README.txt}{here}. This
file outlines the requirements for, and configuration and usage of, the JAR
scripts. The JAR scripts, source code, and required libraries are located in the
same folder as the README file.

@@

An AHP pairwise comparison between the short list packages with respect to
usability and modifiability survey results is conducted by the domain
expert. The tool can be found here. The usability and modifiability survey can
be found
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Experiments.pdf}
{here}. This step is intended to start a conversation about the features of
specific software packages.

@@

\appendix

\section{Appendix}

\subsection{Survey for the Selected Projects}

\ad{Several questions are borrowed from
  \href{https://gitlab.cas.mcmaster.ca/smiths/pub/-/blob/master/Jegatheesan2016.pdf}{Jegatheesan2016},
  and needed to be cited later.}

\subsubsection{Information about the developers and users}

\begin{enumerate}
\item Interviewees' current position/title? degrees?
\item Interviewees' contribution to/relationship with the software?
\item Length of time the interviewee has been involved with this software?
\item How large is the development group?
\item What is the typical background of a developer?
\item How large is the user group?
\item What is the typical background of a user?
\end{enumerate}

@@

\subsection{Empirical Measures Considerations - Processed Data}

Metrics that can be calculated from the raw data.

@@

\subsection{Empirical Measures Considerations - Tool Tests}

\ad{This section is currently a note of unorganized contents. Most parts will
  beremoved or relocated.}

@@

\subsubsection{git-stats}

\href{https://github.com/tomgi/git_stats}{GitHub repo}

@@

\subsubsection{scc}

\href{https://github.com/boyter/scc}{GitHub repo}

\subsubsection{git-of-theseus}

\href{https://github.com/erikbern/git-of-theseus}{GitHub repo}

@@

\subsubsection{hercules}

\href{https://github.com/src-d/hercules}{GitHub repo}

@@

\subsubsection{gitinspector}

\href{https://github.com/ejwa/gitinspector}{GitHub repo}
